Untitled

### 2023-08-01 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
working my way through supervisors comments

In this paper, we describe:

(i) a framework that specifies MOSHEM standards and provides the tools to implement them; and

(ii) the application of that framework to develop and use an initial set of ready4 modules in a utility mapping study.

>>>>

In this paper, we describe:
(i)	criteria for assessing ethical CHEM implementations;
(ii) a prototype software framework for the ethical implementation of CHEMs; and
(iii)	use of the software framework to develop a computational economic model in youth mental health, with an initial focus on outcome valuation.

### 2023-08-01 Markdown/Manuscript/Child_RMDs/MS_Discussion.Rmd
working my way through supervisors comments

Ethical practice is a core expectation of health researchers and computational methods underpin most quantitive research, yet an understanding of what constitutes ethical computational modelling practice in health is underdeveloped [@10.3389/fpubh.2017.00068]. The modeller responsibilities, enabling model attributes and model implementation assessment criteria that we propose can help address this gap.

@@

Currently, many if not most existing CHEMs are insufficiently transparent [@Jalali2021; @McManus2019; @Bermejo2017; @Ghabri2019], reusable [@Feenstra2022; @Emerson2019] and updatable [@Sampson2017; @10.3389/fpubh.2022.899874]]. There appears to be in-principle support from health economists to address these practice shortfalls through greater use of for open-source approaches [@Pouwels2022] that are currently rarely implemented [@Feenstra2022; @Emerson2019]. Existing incentive structures for health economists generally do not promote facilitating peers to reuse their work. Currently, it takes “an extraordinary amount of idealism” to dedicate the substantial time and resources required to author, test, document and maintain even fragile prototype research software that could instead be used to write scientific manuscripts [@anzt2020environment].

>>>>

<!-- Talk about DARTH (https://doi.org/10.1007/s40273-019-00837-x )/ BCEA -->
Ethical practice is a core expectation of health researchers and computational methods underpin most quantitive research, yet an understanding of what constitutes ethical computational modelling practice in health is underdeveloped [@10.3389/fpubh.2017.00068]. The modeller responsibilities, enabling model attributes and model implementation assessment criteria that we propose can help address this gap.

@@

Currently, many if not most existing CHEMs are insufficiently transparent [@Jalali2021; @McManus2019; @Bermejo2017; @Ghabri2019], reusable [@Feenstra2022; @Emerson2019] and updatable [@Sampson2017; @10.3389/fpubh.2022.899874]]. <!-- DELETION --> Existing incentive structures for health economists generally do not promote facilitating peers to reuse their work. Currently, it takes “an extraordinary amount of idealism” to dedicate the substantial time and resources required to author, test, document and maintain even fragile prototype research software that could instead be used to write scientific manuscripts [@anzt2020environment].

### 2023-08-01 Markdown/Manuscript/Child_RMDs/MS_Imp_Guidelines.Rmd
working my way through supervisors comments

We considered prior literature on modelling practice, our own professional experience and the needs of our project, to identify three core ethical responsibilities of CHEM developers, three attributes of CHEMs that enable fulfilment of these responsibilities and six criteria (two for each model attribute) for assessing ethical CHEM implementations.

## Ethical responsibilities of CHEM developers 
We believe that health economists have ethical responsibilities relating to the social acceptability, adequacy for purpose and beneficial impact of their computational models.

Misalignment between the values of model developers and those of the population groups affected by decisions based on their models presents significant ethical risks [@thompson2022escape; @duckett2022journey]. The value judgments of model developers influence the assumptions, selection of model features and standards for evidence that shape the CHEM development process [@HARVARD2020112975]. These value judgments are rarely made explicit, omissions that may lead to socially unacceptable policy recommendations [@duckett2022journey]. For example, to reduce the risk of inequitable policy implementation, it may be important for of a model to predict the benefits of harms of an intervention for different sub-populations [@10.3389/fpubh.2017.00068], but model developers may prefer to allocate a project budget to implementing other model features [@thompson2022escape].

Health economists have duties both to take sufficient care that a CHEM is adequate for the explicit purpose for which it was developed and to provide potential third party users with the means of assessing its adequacy for their proposed purposes [@Erdemir2020; @Feenstra2022; @thompson2019escape; @10.3389/fpubh.2017.00068]. However, it is common for CHEMs to have serious methodological flaws [@carletto_zanuzzi_sammarco_russo_2020; @WONDER2015467]; insufficient validation [@Ghabri2019; @kolovos2017model; @haji2013model], poor reproducibility [@Jalali2021; @McManus2019; @Bermejo2017]; and undeclared errors [@Radeva2020].

Even an acceptable and adequate CHEM will have limited beneficial impact if it not much used, if it is mis-used or when its acceptability and adequacy rapidly decay. The scientific goals of generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39] are advanced when models are appropriately reused in new contexts. Reuse of models as components of other models can also make model implementation more efficient [@Arnold2010; @garcia2021cost].

Common barriers to model re-use include commercial and legal considerations [@Feenstra2022; @Pouwels2022], as well as challenges related to model transferability across jurisdictions [@garcia2021cost]. The temporal window for valid application of CHEMs is often limited by implementation choices that rarely facilitate routine updates [@Sampson2017]. Without ongoing maintenance, a model may become less reliable with time [@garcia2021cost], deterioration that model users may be unaware of, and has a growing risk of being deployed for purposes for which it is poorly suited [@calder2018computational].

##	Criteria for assessing ethical CHEM implementation
The acceptability, adequacy and public benefit responsibilities for CHEM developers are easier to state than to measure.  It may therefore be pragmatic for assessment criteria for ethical modelling practice to instead be based on measurable attributes of the models themselves. As described in Table \@ref(tab:timelygls),  we believe that implementing CHEMs that are transparent, reusable and updatable (TRU) can enable modellers to meet their ethical obligations. We therefore selected these model attributes to use as the basis for deriving ethical assessment criteria.

```{r, child = child_docs_ls$one, echo=FALSE}
```

Transparency has been recommended as a core criterion for assessing ethical public health modelling practice [@10.3389/fpubh.2017.00068]. Guidance on transparency in health economic modelling recommended that model code and data should be clearly documented, potentially with different versions for technical and non-technical users [@Eddy2012]. Notably, the same guidelines, published over ten years ago, did not include recommendations on sharing model code and data. However, more recent guidance recommends publicly dissemination of healthcare model artefacts using online repository services [@Erdemir2020]. Repositories such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection. An essential component of quality assuring health economic models is verification - ensuring that calculations are correct and consistent with model specifications [@techver2019]. The extensiveness of verification checks in models implemented as software projects can be reported using the concept of code coverage [@ERICWONG2010188] - the proportion of model code that has been explicitly tested. Tests should ideally combine both unit tests (to verify that small, isolated sections of code produce the correct output when run independently) and acceptance tests (to verify that the correct output is produced when multiple code components are run together to perform tasks that meet core user-requirements [@martin2003agile]). The nature and extent of individual model authorship contributions can become unclear when models are implemented over longer time-frames with a large and changing group of collaborators [@thompson2022escape]. This issue can be addressed by use of online repository services such as GitHub [@github2007], that provide citation tools and can transparently record all individual code contributions to a modelling project over its lifecycle.

Assessment criteria for a transparent CHEM:

@@

-	T2: It is easy to see who developed and tested each part of the model and to identify the modelling team’s assumptions, judgments and theories about model development and use.

Making a CHEM’s code, data and documentation publicly available is helpful but insufficient for promoting model re-use. The choices that CHEM developers make about model implementation and licensing will also shape who can use a model and for what purposes. Using open-source development platforms and licenses can aid both generalizability and transferability. Compared to using commercial modelling software, authoring CHEMs in an open-source language like R [@RCORE2022] makes it easier to store model algorithms and data in distinct files and locations (as opposed to hard coding - embedding data such as parameter values into source code) which facilitates selective modification of model components. This benefit can be further enhanced if model developers adopt a modular approach, in which a model is constructed from multiple reusable and replaceable sub-models (modules) [@pan2021modular]. To grant permissions to others to use, test and adapt models and their components, health economists can avail of two broad categories of open source licensing options. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. An alternative approach is to use copyleft licenses [@copyleft2022] that can require content users to distribute any derivative works they create under similar open-source arrangements.

Assessment criteria for a reusable CHEM:

>>>>

We considered prior literature on modelling practice, our own professional experience and the needs of our youth mental health modelling project, to identify core ethical responsibilities of CHEM developers, attributes of CHEMs that support fulfilment of these responsibilities and criteria against which these attributes can be assessed.

## Ethical responsibilities of CHEM developers 
Health economists have ethical responsibilities relating to the social acceptability, adequacy for purpose and beneficial impact of CHEMs.

Misalignment between the values of model developers and those of the population groups affected by decisions based on their models presents significant ethical risks [@thompson2022escape; @duckett2022journey]. The value judgments of model developers influence the assumptions, selection of model features and standards for evidence that shape the model development process [@HARVARD2020112975]. These value judgments are rarely made explicit, omissions that may lead to socially unacceptable policy recommendations [@duckett2022journey]. Key choices about the computational implementation of health economic models depend on these value judgments (see Table \@ref(tab:proscons)) and these choices in turn influence the extent to which model implementations are readily adaptable by third parties with alternative value judgments.
<!-- For example, to reduce the risk of inequitable policy implementation, it may be important for of a model to predict the benefits of harms of an intervention for different sub-populations [@10.3389/fpubh.2017.00068], but model developers may prefer to allocate a project budget to implementing other model features [@thompson2022escape]. -->

Model developers have duties both to take sufficient care that a model is adequate for the explicit purpose for which it was developed and to provide potential third party users with the means of assessing its adequacy for their proposed purposes [@Erdemir2020; @Feenstra2022; @thompson2019escape; @10.3389/fpubh.2017.00068]. It is common for health economic models to have serious methodological flaws [@carletto_zanuzzi_sammarco_russo_2020; @WONDER2015467]; insufficient validation [@Ghabri2019; @kolovos2017model; @haji2013model], poor reproducibility [@Jalali2021; @McManus2019; @Bermejo2017]; and undeclared errors [@Radeva2020]. Appropriate computational implementation choices can help address many of these shortcomings, for example by automating quality assurance checks and facilitating manual reviews by third parties.

A health economic model will have limited public benefit if it not much used, if it is mis-used or when its acceptability and adequacy rapidly decay. The scientific goals of generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39] are advanced when models are appropriately reused in new contexts. Reuse of models as components of other models can also make model implementation more efficient [@Arnold2010; @garcia2021cost] but the reusability of models is significantly shaped by computational implementation choices such as how model code and data are authored, documented, licensed and disseminated. Currently, health economic models are rarely implemented computationally in a manner that facilitates routine updates [@Sampson2017], thus limiting the temporal window within which a CHEM can be validly applied. Without ongoing maintenance, a CHEM risks becoming less reliable with time [@garcia2021cost] and is at risk of being deployed for purposes for which it is poorly suited [@calder2018computational].
<!-- Common barriers to model re-use include commercial and legal considerations [@Feenstra2022; @Pouwels2022], as well as challenges related to model transferability across jurisdictions [@garcia2021cost]. -->

##	Sssessing ethical CHEM implementations
The acceptability, adequacy and public benefit responsibilities of health economic modellers are easier to state than to measure.  We have therefore focused our assessment criteria for ethical modelling practice on measurable attributes of computational model implementations. As described in Table \@ref(tab:timelygls),  we believe that implementing CHEMs that are transparent, reusable and updatable (TRU) can enable modellers to meet their ethical obligations. We therefore selected these model attributes to use as the basis for deriving ethical assessment criteria.

```{r, child = child_docs_ls$two, echo=FALSE}
```

**Transparency** has been recommended as a core criterion for assessing ethical public health modelling practice [@10.3389/fpubh.2017.00068]. Guidance on transparency in health economic modelling recommended that model code and data should be clearly documented, potentially with different versions for technical and non-technical users [@Eddy2012]. Notably, the same guidelines, published over ten years ago, did not include recommendations on sharing model code and data. However, more recent guidance recommends publicly dissemination of healthcare model artefacts using online repository services [@Erdemir2020]. Repositories such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection. An essential component of quality assuring health economic models is verification - ensuring that calculations are correct and consistent with model specifications [@techver2019]. The extensiveness of verification checks in models implemented as software projects can be reported using the concept of code coverage [@ERICWONG2010188] - the proportion of model code that has been explicitly tested. Tests should ideally combine both unit tests (to verify that small, isolated sections of code produce the correct output when run independently) and acceptance tests (to verify that the correct output is produced when multiple code components are run together to perform tasks that meet core user-requirements [@martin2003agile]). The nature and extent of individual model authorship contributions can become unclear when models are implemented over longer time-frames with a large and changing group of collaborators [@thompson2022escape]. This issue can be addressed by use of online repository services such as GitHub [@github2007], that provide citation tools and can transparently record all individual code contributions to a modelling project over its lifecycle.

The **transparency** of a CHEM implementation can be assessed against the criteria:

@@

-	T2: It is easy to see who developed and tested each part of the CHEM and to identify the modelling team’s assumptions, judgments and theories about CHEM development and use.

Model **reusability** is the second attribute that we identified as being core to ethical CHEM implementations. Making a CHEM’s code, data and documentation publicly available is now considered “gold-standard” practice but is not sufficient to promote model re-use. Computational implementations that store model algorithms and data in distinct files and locations (as opposed to embedding data such as parameter values into source code) facilitate selective modification of model components. Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task), has been recommended as good practice for scientific computing [@Wilson_2017] and can facilitate model generalisability. The concept of inheritance [@8717448], a feature of object-oriented programming, facilitates selective editing of algorithms when transferring models to different decision contexts [@ready4oop2022]. This benefit can be further enhanced if model developers adopt a modular approach, in which a model is constructed from multiple reusable and replaceable sub-models (modules) [@pan2021modular]. To grant permissions to others to use, test and adapt models and their components, health economists can avail of two broad categories of open source licensing options. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. An alternative approach is to use copyleft licenses [@copyleft2022] that can require content users to distribute any derivative works they create under similar open-source arrangements. 
<!-- Compared to using commercial modelling software, authoring CHEMs in an open-source language like R [@RCORE2022] makes it easier to -->
<!-- Using open-source development platforms and licenses can aid both generalizability and transferability. -->
<!-- The choices that CHEM developers make about model implementation and licensing will also shape who can use a model and for what purposes.  -->
The **reusability** of a CHEM implementation can be assessed against the criteria:

### 2023-08-01 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd
working my way through supervisors comments

The ready4pack library depends on two other module authoring libraries. Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task), has been recommended as good practice for scientific computing [@Wilson_2017].  The ready4fun library contains tools for authoring functions in a consistent house style that automatically generates basic documentation for each function. Functions to implement model algorithms can be associated with a module via a special type of function called a method. Tools from the ready4class library can help streamline and standardise the authoring of module data structures and their associated methods and to automatically generate basic documentation for each module.

@@

```{r, child = child_docs_ls$two, echo=FALSE}
```

>>>>

The ready4pack library depends on two other module authoring libraries. <!-- MOVED CONTENT -->The ready4fun library contains tools for authoring functions in a consistent house style that automatically generates basic documentation for each function. Functions to implement model algorithms can be associated with a module via a special type of function called a method. Tools from the ready4class library can help streamline and standardise the authoring of module data structures and their associated methods and to automatically generate basic documentation for each module.

@@

```{r, child = child_docs_ls$three, echo=FALSE}
```

### 2023-08-01 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
working my way through supervisors comments

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four].  In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. These representations can be described in words and pictures (a conceptual model), in equations (a mathematical model) or in computer code (a computational model). The predictions reported in health economic studies are typically generated by the execution of a computer program which applies a computational model to compatible data inputs (e.g., parameter values) and performs a sequence of numeric calculations.

Computational models are now widely used to inform health policy and system design [@dakin2015influence; @Erdemir2020; @LO2022100570; @Christene006827]. This level of influence has concomitant ethical responsibilities for model developers that are often poorly understood and inadequately fulfilled [@10.3389/fpubh.2017.00068; @pliakos2021ethics; @thompson2022escape; @calder2018computational].

Computational health economic models (CHEMs) can be implemented using specialized commercial software or authored as bespoke software projects in a programming language. Advantage of commercial modelling tools are simplicity and ease of use, but a software development approach facilitates development of models that may be more realistic, transparent, reusable and adaptable [@incerti2019r]. For CHEMs that are implemented as software projects, a major early decision is selection of an appropriate software framework. A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. Advantages of using software frameworks include facilitating code reuse and extension, promoting good programming practice and the capability to provide enhanced functionality and performance without additional effort by developers [@edwin2014software]. However, software frameworks can be challenging and time consuming to create [@edwin2014software] and then difficult for others to learn, often requiring developers to undergo specialist training [@myllarniemi2018development]. There is also a risk that a software framework may become excessively complex over time [@edwin2014software].

We are developing a model to explore multiple economic questions relating to the mental health of young people aged 12 to 25.  When making choices about how to implement this model computationally, we wished to facilitate ethical development and use. However, we are not aware of any software framework for implementing CHEMs that adhere to explicitly stated ethical requirements.

>>>>

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four].  In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. A health economic model should be capable of being described using three isomorphic representational approaches - words and figures (a conceptual model), equations (a mathematical model) and computer code (a computational model). Health economic scientific manuscripts typically focus on describing conceptual and mathematical model representations. However, the predictions reported in health economic studies are now almost always generated by the execution of a computer program that applies a computational model to compatible data inputs (e.g., parameter values) and performs a sequence of numeric calculations. Although health economists report strong interest in making the computational implementation of their models more transparent and accessible [@Pouwels2022], open-source model implementations remain rare [@Feenstra2022; @Emerson2019].

Computational health economic models (CHEMs) can be implemented using specialized commercial software or authored as software development projects in a programming language such as Python [@python2009] or R [@RCORE202]. Advantages of commercial modelling tools are simplicity and ease of use, but a software development approach may facilitate development of computational models that are more realistic, transparent, reusable and adaptable [@incerti2019r]. Which approach is most appropriate for a health economic modelling project to adopt may depend on both modelling team value judgments and the resource and time constraints within which projects need to be implemented (see Table \@ref(tab:proscons)).

```{r, child = child_docs_ls$one, echo=FALSE}
```
<!-- on considerations such as model comprehensiveness, generalisability, reproducibility and transferability -->
For CHEMs that are implemented as software development projects, a major early decision is the selection of an appropriate software framework. A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. A software framework provides a foundation for developing software applications with shared resources, e.g., algorithms and data files, that can be modified to suit specific needs. Software frameworks have widely been developed and implemented in computer science, for example, PyTorch [@PyTorch2019] for machine learning using Python.

Advantages of using software frameworks include facilitating code reuse and extension, promoting good programming practice and the capability to provide enhanced functionality and performance without additional effort by developers [@edwin2014software]. However, software frameworks can be challenging and time consuming to create [@edwin2014software] and then difficult for others to learn, often requiring model developers to undergo specialist training [@myllarniemi2018development]. There is also a risk that a software framework may become excessively complex over time [@edwin2014software].

<!-- We are developing a model to explore multiple economic questions relating to the mental health of young people aged 12 to 25.   -->
<!-- When making choices about how to implement this health economic model computationally, we wished to facilitate ethical development and use. -->
Computational models are now widely used to inform health policy and system design [@dakin2015influence; @Erdemir2020; @LO2022100570; @Christene006827]. This level of influence has concomitant ethical responsibilities for model developers that are often poorly understood and inadequately fulfilled [@10.3389/fpubh.2017.00068; @pliakos2021ethics; @thompson2022escape; @calder2018computational]. There is limited guidance on software frameworks for implementing CHEMs that adhere to explicitly stated ethical requirements.

### 2023-08-01 Markdown/Manuscript/Child_RMDs/MS_Tables.Rmd
working my way through supervisors comments

```{r echo=FALSE}
if(params$X@tables_in_body_lgl){
  child_docs_ls <- list(one = "../Child_RMDs/MS_Tables_1.RMD",
                        two = "../Child_RMDs/MS_Tables_2.RMD",
                        three = "../Child_RMDs/MS_Tables_3.RMD")
}else{
    child_docs_ls <- list(one = NULL,
                      two = NULL)
}
```

>>>>

```{r echo=FALSE}
if(params$X@tables_in_body_lgl){
  child_docs_ls <- list(one = "../Child_RMDs/MS_Tables_1.RMD",
                        two = "../Child_RMDs/MS_Tables_2.RMD",
                        three = "../Child_RMDs/MS_Tables_3.RMD",
                        four = "../Child_RMDs/MS_Tables_4.RMD")
}else{
    child_docs_ls <- list(one = NULL,
                      two = NULL,
                      three = NULL,
                      four = NULL)
}
```

### 2023-08-01 Markdown/Manuscript/Child_RMDs/MS_Tables_2.Rmd
working my way through supervisors comments

```{r cpkgs, tab.cap='Modelling toolkit R packages for developing and using MOSHEMs that meet framework standards', tab.id = 'cpkgs', echo = F, results = 'asis'}

@@

df <- tibble::tibble(Package = c("ready4",
                                       "ready4fun",
                                       "ready4class",
                                       "ready4pack",
                                       "ready4use",
                                       "ready4show"),
                           # Version = Package %>%
                           #   purrr::map_chr(~utils::packageDescription(.x) %>%
                           #                    purrr::pluck("Version")),
                           Focus = c("Foundation",
                                     "Module algorithms",
                                     "Module structures",
                                     "Module libraries",
                                     "Datasets",
                                     "Analyses"),
                           A = c("3","2-3","2","1-2,5,7","1-2,7","4,6"), 
                           R = c("4","5","4-5","1,3,6","3,7",""),
                           U = c("","","","1-3","1",""),
                           # E = c("", "1","1-2","1","3",""),
                           # L = c("","","","1,3","2-3",""),
                           # Y = c("","","","","","1-2"),
                           `Depends on` = Package %>%
                             purrr::map(~ {
                               utils::packageDescription(.x) %>%
                                 `[`(c("Depends", "Imports")) %>%
                                 purrr::map(~{
                                   if(is.null(.x)){
                                     character(0)

@@

                                       .x %>%
                                         strsplit(",\\n") %>%
                                         purrr::flatten_chr() %>%
                                         purrr::map(~strsplit(.x,", ") %>%
                                                      purrr::flatten_chr()) %>%
                                         purrr::flatten_chr() %>% sort() %>%
                                         purrr::discard(~startsWith(.x,"R "))
                                       }
                                   }) %>%
                                 purrr::flatten_chr() %>%
                                 unique() %>%
                                 sort() %>%
                                 purrr::map_chr(~{
                                   updated_1L_chr <- stringr::str_replace_all(.x,"\\n"," ")
                                   problem_idx_1L_chr <- stringr::str_locate(updated_1L_chr," ")[1,1] %>%
                                     unname()
                                   if(!is.na(problem_idx_1L_chr))
                                     updated_1L_chr <- updated_1L_chr %>%
                                     stringr::str_sub(end = problem_idx_1L_chr-1)
                                   updated_1L_chr %>% 
                                     trimws(which = "left") 
                                   })
                               }) %>% 
                             purrr::map_chr (~ .x %>% paste0(collapse = " "))) %>% 
  as.data.frame() 
    df$Ref <- c("@ready42022",
                    "@ready4fun2022",
                    "@ready4class2022",
                    "@ready4pack2022",
                    "@ready4use2022",
                    "@ready4show2022")
df <- df %>%
  dplyr::mutate(Standard = purrr::pmap_chr(df,
                                           ~paste0("A",
                                                  ..3,
                                                  paste0(rep("&nbsp;",
                                                      2+max(nchar(df[,3]))-nchar(..3)),
                                                      collapse = ""#
                                                  ifelse(..4=="","","R"),
                                                  ..4,
                                                  paste0(rep("&nbsp;",
                                                      2+max(nchar(df[,4]))-nchar(..4)),
                                                      collapse = ""#
                                                  ifelse(..5=="","","U"),
                                                  ..5,
                                                  paste0(rep("&nbsp;",
                                                      1+max(nchar(df[,5]))-nchar(..5)),
                                                      collapse = ""#
                                                      )#,
                                                  # ifelse(..6=="","","E"),
                                                  # ..6,
                                                  # paste0(rep("&nbsp;",
                                                  #     1+max(nchar(df[,6]))-nchar(..6)),
                                                  #     collapse = ""#
                                                  #     ),
                                                  # ifelse(..7=="","","L"),
                                                  # ..7,
                                                  # # paste0(rep("&nbsp;",
                                                  # #     1+max(nchar(df[,7]))-nchar(..7)),
                                                  # #     collapse = ""#
                                                  # #     ),
                                                  # ifelse(..8=="","","&nbsp; &nbsp; &nbsp; &nbsp; Y"),
                                                  # ..8,
                                                  # paste0(rep("&nbsp;",
                                                  #     1+max(nchar(df[,8]))-nchar(..8)),
                                                  #     collapse = ""#
                                                  #     )
                                                  ))) %>%

  dplyr::select(Package,
                Ref,
                Focus,
                Standard,
                `Depends on`)

names_chr <- names(df)
df <- data.frame(df[,1],
                 rep(ifelse(params$X@outp_formats_chr[1] == "Word"," ","&nbsp; &nbsp; &nbsp;"),
                     nrow(df)),
                 df[,2:3],
                 rep(ifelse(params$X@outp_formats_chr[1] == "Word"," ","&nbsp; &nbsp; &nbsp;")
                     ,nrow(df)),
                 df[,4],
                 #rep("&nbsp; &nbsp; &nbsp;",nrow(df)),
                 df[,5])
colnames(df) <- c("Package",
                  ifelse(params$X@outp_formats_chr[1] == "Word","XXX"," "),
                  "Ref", "Focus", 
                  ifelse(params$X@outp_formats_chr[1] == "Word","YYY"," "),
                  "Standard     ",# " ",
                  "Depends on these R libraries")
#df$Package <- "![](../Data/images/ready4.png){ width=10% }"#"\\includegraphics[valign=T,scale=2.5,raise=2mm]{../Data/images/ready4.png}"
df$`Standard     `  <- df$`Standard     ` %>% stringr::str_replace_all("&nbsp;"," ")
if(params$X@outp_formats_chr[1] == "Word") {
  library("flextable")
  library("ftExtra")
  df[,c(1,3:4,6:7)] %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    df$Ref<-NULL
    #df$Package <- ""
    df$Package  <- df$Package %>% 
                      purrr::map_chr(~paste0("\\includegraphics[valign=M,width=1.9cm,raise=4mm]{../Data/images/",.x,"_logo.png}"))#scale=0.4,raise=3mm]
    # df$Package <- "\\includegraphics[valign=M,scale=0.08,raise=3mm]{../Data/images/ready4.png}"
    df[,c(1,3,5:6)]   %>%
    # kableExtra::kbl(format = "markdown",
    #                   escape = F,
    #                 caption = knitr::opts_current$get("tab.cap"))
      kableExtra::kbl(booktabs = T, 
                      escape = F,
                      caption = knitr::opts_current$get("tab.cap")) %>%
      kableExtra::kable_paper(full_width = F) %>%
      kableExtra::column_spec(3,latex_valign = "m") %>%
      # kableExtra::column_spec(1, 
      #                         image = kableExtra::spec_image(rep("../Data/images/ready4.png",nrow(df)), 
      #                                                           150, 150)) %>%
      kableExtra::column_spec(4, width = "24em",latex_valign = "m")
    }
```

>>>>

```{r timelygls, tab.cap='Framework standards', tab.id = 'timelygls', echo = F, results = 'asis'}

@@

df <- data.frame(`Objective` = rep("",20),
  # `Models are` = c(rep("Transparent",6),
  #                                 rep("Iterative",4),
  #                                 rep("Modular",2),
  #                                 rep("Epitomising",3),
  #                                 rep("Licensed",3),
  #                                 rep("Yielding",2)),
  #                Goal = c(rep("people can easily see how a model has been implemented and tested",6),
  #                         rep("a model is routinely updated to maintain and improve validity",4),
  #                         rep("models and their components can be combined to extend their scope",2),
  #                         rep("model code is sufficiently general to be be re-used in multiple decision contexts",3),
  #                         rep("a model, its components and derivatives are persistently re-usable by other modeller",3),
  #                         rep("a model can be simply, flexibly and reliably used as a decision aid",2)),
                 Standard = c(paste0("A",#"T", 
                                      1:7
                               paste0("R",1:9),# paste0("M",1:2),
                               # paste0("E",1:3),
                               # paste0("L",1:3),
                               # paste0("Y",1:2),
                               paste0("U",#"I",
                                      1:4)
                 Meaning = c(
                   "Open online repositories are used to permanently archive, uniquely identify and transparently record authorship and development history of model code and data.",#T1
                   # T2 MISSING - Merged With T1
                   "Model code and data are documented.",#<T2, Was T3
                   "Model code uses consistent and abstracted syntax.",# T3, Was T4 -->
                   "Literate programming is used to implement model analyses.",#T4,Was T5
                   "Model code coverage is reported.", #T5, New, replaces version now moved to T4 
                   "All parts of a study analysis and reporting workflow can be reproduced and/or replicated.", #T6, was T7, Replaces Deleted T6 
                   "Model code and data are distributed with tools to support appropriate citation.",#*L3: New, new location
                   "Model code is made available for re-use under copyleft or permissive licenses.",#L1: **
                   "Non-confidential model data is licensed for liberal re-use (potentially subject to ethical use terms).",#L2
                   "Model code and data are stored and managed separately.",#*M1: Modified (lost ref to licensed and cited) 
                   "Model code uses encapsulation and inheritence for data structures.",#*M2: Replacement - Old version now merged with E1 # And with E2: "Modules are inheriting data-structures.",#E2
                   "Model code uses functions to implement algorithms.",#E1: Modified (lost ref to OOP and merged with M2) -->
                   "Model code is distributed as code libraries.",
                   "Test data is available to demonstrate the transferability of model code.",#E3: **
                   # OLD CITATION
                   "Statistical models are distributed with tools for making out of sample predictions.",#*Y1: New, Replaces Old Y1 (now Y2)
                   "User-interfaces allow non-technical users to configure and use models.",#Y2: Was Y1     
                   "Model code and data are version controlled.",#*I1, Was I2 
                   "The signficiance and status of code and data updates are indicated with semantic versioning and release types.",#I2, Was I1 
                   "Continuous integration is used to verify model code updates.",#I3 
                   "Deprecation conventions are used to retire model code and data."#**I4 
                 check.names = F)
if(params$X@outp_formats_chr[1] == "Word") {
  library("flextable")
  library("ftExtra")
  df %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()

@@

    df %>%
      kableExtra::kbl(booktabs = T, 
                      escape = F,
                      caption = knitr::opts_current$get("tab.cap")) %>%
      kableExtra::kable_paper(full_width = F) %>%
      kableExtra::column_spec(1, bold=T) %>%
      # kableExtra::collapse_rows(columns = 1) %>%
      kableExtra::column_spec(3, width = "35em") %>%
      kableExtra::pack_rows("Accountable - it is easy to see who developed, tested and applied a model and how they did it.",1,7) %>%
      kableExtra::pack_rows("Updatable - a model and its components are maintained and continuously improved.",17,20) %>%
      # kableExtra::pack_rows("Modular - models and their components can be combined to extend their scope",11,12) %>%
      # kableExtra::pack_rows("Epitomising - model code can be re-used in other decision contexts",13,15) %>%
      kableExtra::pack_rows("Reusable - a model and its components can be used in other models and by other modellers.",8,16) #%>%
      # kableExtra::pack_rows("Yielding - a model can be simply, flexibly and reliably used as a decision aid",19,20)

    }
```

### 2023-08-01 Markdown/Manuscript/Child_RMDs/MS_Tables_4.Rmd
working my way through supervisors comments

                                           "All module data-structures and algorithms are distributed as five code libraries [@hamilton_matthew_2022_6084467; @hamilton_matthew_2022_6084824; @hamilton_matthew_2022_6116701; @gao_caroline_2022_6130155; @matthew_p_hamilton_2021_5646669].",

>>>>

                                           "All module data-structures and algorithms are distributed as code libraries [@hamilton_matthew_2022_6084467; @hamilton_matthew_2022_6084824; @hamilton_matthew_2022_6116701; @gao_caroline_2022_6130155; @matthew_p_hamilton_2021_5646669].",
